What to simulate? Inferring the right direction for mental rotation

### 2014-02-04 man/cogsci-2014/cogsci-2014.tex
Misc wording changes

One of our most astonishing cognitive feats is the ability to
envision, manipulate, and plan with objects -- all without actually
perceiving them. This capacity for ``mental simulation'' has been
widely studied and has sparked intense debate about the underlying
representation of mental images \cite<e.g.,>{Kosslyn:2009tj,
  Pylyshyn:2002vk}. But this debate hasn't addressed one of the most
fundamental questions about mental simulation: how people decide what
they should simulate.

Mental rotation provides a simple example of the decision problem
posed by simulation.  In the classic experiment by
\citeA{Shepard1971}, participants viewed images of three-dimensional
objects and had to determine whether the images depicted the same
object (which differed by a rotation) or two separate objects (which
differed by a reflection and a rotation). The result was that people's
response times had a strong linear correlation with the minimum angle
of rotation, leading to the conclusion that people solve this task by
``mentally rotating'' the objects until they are congruent.  This
conclusion leaves several questions unanswered, however. How do people
know the axis around which to rotate the objects? If the axis is
known, how do people know which direction to rotate the objects?  And
finally, how do people know how long to rotate?

@@

Previous models of mental rotation have, for the most part, focused on
the representation of mental images, rather than how people decide
\textit{which} mental images to generate. \citeA{Kosslyn:1977tv}
proposed a model of the mental imagery buffer, but not the planning
process of how it is used. Similarly, \citeA{Julstrom:1985va} and
\citeA{Glasgow:1992tj} were mostly concerned with modeling the
representational format underlying imagery. Although
\citeA{Anderson1978} emphasized the importance of considering both
representation and process, he dismissed the problem of determining
the direction of rotation as a ``technical difficulty''.

@@

If people are not computing the rotation beforehand, what might they
be doing? To answer this question, we perform a ``rational analysis''
of the problem of mental rotation
\cite{Marr:1983to,anderson90,Shepard:1987tt}. At the computational
level, we could say that the \textit{problem} is to determine which
spatial transformations an object has undergone based on two
depictions of that object which do not include information about point
correspondences across images. At the algorithmic level, we are
constrained by the notion that mental images must be transformed in an
analog manner (or in a way that is at least approximately analog), and
that mental images are time-consuming and effortful to generate. Thus,
the \textit{goal} is to make this determination while performing a
minimum amount of computation (i.e., as few rotations as possible).

The original ``congruency'' hypothesis \cite{Shepard1971} actually is
a rational solution to this problem, in the sense that the smallest
amount of computation coincides with rotating through the minimum
angle. However, it violates the constraint that we do not know the
points of correspondence between the images, which is exactly what
necessitates the use of imagery. Noting that a rational solution need
not necessarily maintain a single trajectory of rotation, we explore
an alternative rational model, which--rather than computing the angle
of rotation ahead of time--engages in an ``active sampling'' strategy.

Active sampling is the idea that, when given information is
incomplete, people gather new information in a manner that increases
certainty about the problem space. This notion has recently gained
support in several areas of cognitive science
\cite<e.g.,>{Nelson2007,Gureckis:2012gu}, including other spatial
domains \cite{Juni:2011vo}. In the case of mental rotation, if people
do not know the minimum angle beforehand, then actively performing
rotations may be the best way to gather evidence about the similarity
between the observed shapes.

@@

In this section, we formalize our rational analysis, and propose four
models of mental rotation: one based on existing models; two which are
extensions of the first but with relaxed assumptions; and one based on
the active sampling approach.

@@

Formally, we denote the shapes as $X_a$ and $X_b$ and assume $X_b$ is
generated by a transformation of $X_a$, i.e. $X_b=f(X_a, \theta, h)$,
where $\theta$ is a rotation, $\hi$ is the hypothesis that the images
depict the same object, and $\hf$ is the hypothesis that the images
depict mirror-image objects. The posterior probability of each
hypothesis given the observed shapes is then:
\begin{equation}
  p(h\ \vert\ X_a, X_b) \propto \int p(X_b\ \vert\ X_a, \theta, h)p(X_a)p(h)p(\theta)\dif\theta,
  \label{eq:posterior}
\end{equation}
where $p(X_b\ \vert\ X_a, \theta, h)$ is the probability that $X_b$
was generated from $X_a$, and $p(X_a)$ is the probability of observing
$X_a$. Because we ultimately want to determine which hypothesis is
more likely, the quantity of interest is a ratio $\ell:=p(\hi\ \vert\
X_a, X_b) / p(\hf\ \vert\ X_a, X_b)$ which (assuming that all
rotations are equally likely) simplifies to:
\begin{equation}
  \ell = \frac{\left(\int p(X_b\ \vert\ X_a, \theta, \hi)\dif\theta\right)p(\hi)}{\left(\int p(X_b\ \vert\ X_a, \theta, \hf)\dif\theta\right)p(\hf)}.
  \label{eq:lh-ratio}
\end{equation}
If $\ell > 1$, then we accept the hypothesis that the images depict
the same object ($\hi$); if $\ell < 1$, then we accept the hypothesis
that the images depict flipped objects ($\hf$).

@@

\begin{figure*}[t]
  \begin{center}
    \includegraphics[width=\textwidth]{../../figures/D/model-traces.pdf}
    \caption{\textbf{Example stimuli and model behavior.} On the left
      are two stimulus pairs, one ``same'' (top) and one ``flipped''
      (bottom), both with a rotation of $120^\prime$. The grey plots
      show the actions that each model took for these stimuli. The
      color of each point (action) indicates whether the model's
      current mental image was generated under the hypothesis that the
      objects are the same (red) or flipped (blue). The similarity
      functions for each of these hypotheses are shown on the right.}
    \label{fig:model-traces}
  \end{center}
\end{figure*}

@@

One hypothesis in the literature is that people compute the direction
and extent of rotation beforehand using \textit{a prior} knowledge of
the correspondence between points in the images
\cite{Funt:1983wn,Just:1985uu}.  To reflect this hypothesis, we
created an ``oracle'' model which is told which points on each shape
correspond. From that correspondence, it computes the correct rotation
and rotates through it.

@@

An intuitive and simple model which does not know which way to rotate
\textit{a priori} might first rotate in the direction that increases
similarity, and then stop rotating once a ``match'' is found, as
determined by a threshold on the value of $S$. If all rotations have
been exhausted, then flip, and try rotating again. We assume the value
of $S$ which is greater than the threshold corresponds to the true
$\theta$, or a point near the true $\theta$. So, as with the \Oc{}
model, Eq.~\ref{eq:lh-ratio} becomes a likelihood ratio.

@@

In the current formulation of the problem, choosing the threshold is
straightforward because we know both the exact geometry of the shapes
and that a linear transformation exists which will perfectly align
them. However, this choice is not always clear \textit{a priori}, as
the global optimum will depend on many factors (e.g., shape
complexity, dimensionality, perceptual uncertainty, and whether the
shapes are identical).  One way to deal with the problem of choosing a
threshold would simply be to perform an exhaustive search or use a
global optimization strategy. However, these strategies would not
result in the linear response time found by \citeA{Shepard1971}. A
second alternative, which could potentially demonstrate the linear
response time result, might instead just perform a Hill Climbing
(\Hc{}) search for a \textit{local} maximum. As with the \Oc{} and
\Th{} models, Eq.~\ref{eq:lh-ratio} becomes a generalized likelihood
ratio test.

@@

While the previous few models have all focused on \textit{searching}
for the global maximum, the goal is really to just approximate
Eq.~\ref{eq:lh-ratio} while still obeying the constraint of sequential
rotations. Based on evidence that people use ``active sampling''
\cite<e.g.,>{Nelson2007,Gureckis:2012gu}, we hypothesize a model that
actively monitors the usefulness of the rotations it is
performing. Instead of searching for a maximum, we maintain a
probability distribution over our \textit{estimate} of
Eq.~\ref{eq:lh-ratio}, and then sample actions which are expected to
improve that estimate.  This strategy has several benefits. First, it
does not make assumptions about the shape or scale of the similarity
function. Second, by choosing to sample places which are informative,
this method implicitly minimizes the amount of rotation; yet, because
it is constrained to sequential rotations, it should still exhibit the
linear effect of the angle.

Formally, we denote $Z_h$ as our estimate of the likelihood for
hypothesis $h$, and write its distribution as:
\begin{equation}
  p(Z_h) = \int \left[\int S(\Xb, f(\Xa, \theta, h))p(\theta)\dif\theta\right] p(S)\dif S.
\end{equation}

This is a technique known in the machine-learning literature as
\textit{Bayesian Quadrature} (BQ)
\cite{Diaconis:1988uo,OHagan:1991tx,Osborne:2012tm}. Specifically, we
use a Gaussian Process (GP) prior for $p(S)$, as in
\citeA{Osborne:2012tm}.  Denoting $S_h=S(\Xb, f(\Xa, \theta, h))$, we
first place a GP prior on the log of $S_h$ in order to enforce
positivity after it is exponentiated, i.e. $\mathbb{E}[Z_h] \approx
\int \exp(\mu_h(\theta))p(\theta)\dif\theta$, where $\mu_h:=\mu(\log
S_h)$ is the mean of the log-GP.  However, this integral is still
intractable. To approximate it, we fit a second GP over points sampled
from the log-GP, which we denote as
$\bar{S}_h:=\exp(\mu_h)$.\footnote{Note that the approach used here
  differs slightly from the one presented in \citeA{Osborne:2012tm}
  and is based on personal communication with one of the authors
  \cite{Duvenaud:2013td}.} Then, from \citeA{Duvenaud:2013td}:
\begin{align}
  \mathbb{E}[Z_h] &\approx \int \bar{\mu}_h(\theta)p(\theta)\dif\theta,\\
  \mathbb{V}(Z_h) &\approx \iint \mathrm{Cov}_h(\theta,
  \theta^\prime)\bar{\mu}_h(\theta)\bar{\mu}_h(\theta^\prime)p(\theta)p(\theta^\prime)\dif\theta\dif\theta^\prime,
\end{align}
where $\bar{\mu}_h:=\mu(\bar{S}_h)$ is the mean of the second GP, and
$\mathrm{Cov}_h:=\mathrm{Cov}(\log S_h)$ is the covariance of the
log-GP.

@@

We randomly generated 20 shapes of five or six vertices (e.g.,
Fig.~\ref{fig:model-traces}). For each shape, we computed 20 ``same''
and 20 ``flipped'' stimuli pairs, with rotations ($\theta$) spaced at
$20^\prime$ increments between 0 and 360 (with 0 and 180 degree
rotations repeated twice, in order to gather an equal number of
responses for each angle between 0 and 180). ``Same'' pairs were
created by rotating $\Xa$ by $\theta$; ``flipped'' pairs were first
reflected $\Xa$ across the $y$-axis, then rotated by $\theta$.

We generated five additional shapes to be used in a practice block of
10 trials. Across these trials, each shape and each angle (60, 120,
180, 240, or 300) was repeated twice (once ``flipped'' and once
``same'') such that no shape was presented at the same angle twice. We
also generated a sixth shape to be included along with the
instructions.  This shape had both a ``flipped'' and ``same'' version,
each rotated to 320 degrees.

@@

All participants saw the same 10 practice trials, as described
above. There were 720 unique experimental stimuli (20 shapes $\times$
18 angles $\times$ 2 reflections), though because stimuli with
rotations of $0^\prime$ or $180^\prime$ were repeated twice, there
were 800 total experimental stimuli. These stimuli were split across
eight conditions in the following manner: first, stimuli were split
into four blocks of 200 trials. Within each block, each shape was
repeated ten times and each rotation was repeated ten times (five
``same'', five ``flipped''), such that across all blocks, each
stimulus appeared exactly once. Each block was then split in
half. Participants completed two half-blocks (not necessarily both
from the original full block).

@@

Out of the 247 participants, 200 (81\%) were included in our
analyses. Of the rest, we excluded 10 (4\%) because of an experimental
error, 6 (2.4\%) because they had already completed a related
experiment, and 31 (12.6\%) because they failed a comprehension
check. A participant was marked as passing the comprehension check if
they answered at least 85\% of ``easy'' trials correctly. ``Easy''
trials were defined to be those for which the presented shapes
differed by a rotation of no more than 20 degrees. Additionally, we
excluded 82 trials for which the response time was either less than
100 milliseconds or greater than 20 seconds.

For each model, we ran 50 samples for each stimulus that participants
viewed (with the exception of stimuli with rotations of $0^\prime$ or
$180^\prime$, in which case we ran double the number of samples). The
step size parameter ($\sigma_\epsilon$) was fit to human data for each
of the models, resulting in $\sigma_\epsilon=0.6$ for the \Th{} and
\Bq{} models and $\sigma_\epsilon=0.1$ for the \Oc{} and \Hc{}
models. We also ran the models under two different priors,
$p(h=0)=0.5$ (the ``equal'' prior) and $p(h=0)=0.55$ (the ``unequal''
prior). This had an effect on accuracy for the \Hc{} model, and an
effect on both accuracy and response time for the \Bq{} model. Below,
we report results for the \Hc{} model with an unequal prior, and the
\Bq{} model with both priors.

@@

For analyses of response time, confidence intervals around harmonic
means were computed using a bootstrap analysis using 10000 bootstrap
samples (sampled with replacement). Response times were only
considered for correct responses.  We also used a bootstrap analysis
of 10000 bootstrap samples to compute the confidence intervals around
both Spearman ($r_s$) and Pearson ($r$) correlations.  For analyses of
accuracy, confidence intervals were computed from a binomial
proportion with a Jeffrey's beta prior.  To test if participants or a
model was above chance on a particular stimulus, we used the same
binomial proportion, and simply tested whether $p>0.05$ for the 50\%
mark.

@@

The average accuracy across all stimuli was \ExpAccuracy{}, though
there were \ExpNumChance{} individual stimuli (out of 720) for which
people were not above chance.  The minimum angle was also correlated
with participants' average per-stimulus accuracy, as shown in
Fig.~\ref{fig:theta-accuracy-corrs}, though significantly more so for
``same'' pairs. This is the same result found both by
\citeA{Cooper:1975wp} and \citeA{Gardony:2013gn}.

@@

The number of actions taken by the \Th{} model was about as correlated
with the minimum angle as human response times were
(Fig.~\ref{fig:theta-time-corrs}), though this effect was less strong
for individual shapes (e.g.,
Fig.~\ref{fig:response-time-stimulus}). Like the \Oc{} model, the
\Th{} model had 100\% accuracy, and thus there was no correlation
between minimum angle and accuracy. The \Th{} was able to explain a
moderate amount of the variance in human response times, with a
correlation of \ThTimeCorr{} (Fig.~\ref{fig:human-model-corrs}).

As noted above, we fit $\sigma_\epsilon=0.6$ for the \Th{} model. This
has the interesting effect of actually causing the \Th{} to
\textit{over}rotate (e.g., the ``flipped'' pair trace in
Fig.~\ref{fig:model-traces}). This is because the step size is large
enough that it sometimes misses the global maximum, and has to do
another full rotation to find it. However, when we use a lower step
size (so that it does not miss the maximum), the fit to human data is
slightly worse.

@@

The \Bq{} model was quite accurate overall, both with the equal prior
(\BqAccuracy{}) and unequal prior (\BqpAccuracy{}). With the equal
prior, there was \BqNumChance{} stimulus for which it was not above
chance; with the unequal prior, there were \BqpNumChance{}. There was
a significant effect of rotation on the \Bq{} model's accuracy
(Fig.~\ref{fig:theta-accuracy-corrs}), and it explained about the same
amount of variance in people's accuracy as the \Hc{} model: for the
equal prior, \BqAccuracyCorr{}; for the unequal prior,
\BqpAccuracyCorr{}.

Because the \Bq{} model relied on Eq.~\ref{eq:lh-ratio} for its
stopping criteria (as opposed to just finding a maximum), the prior
$p(h)$ had a fairly significant effect, as shown in
Fig.~\ref{fig:response-time-accuracy}. With just a small bias of
$p(h=0)=0.55$, there is a clear separation in response times that is
similar to the separation seen in human response times. The prior also
had an effect on accuracy, though this effect did not reflect the
trends in human behavior.

>>>>

One of the most astonishing cognitive feats is our ability to
envision, manipulate, and plan with objects -- all without actually
perceiving them. This capacity for ``mental simulation'' has been
widely studied, including an intense debate about the underlying
representation of mental images \cite<e.g.,>{Kosslyn:2009tj,
  Pylyshyn:2002vk}. But this debate hasn't addressed one of the most
fundamental questions about mental simulation: how people decide what
they should simulate.

Mental rotation provides a simple example of the decision problem
posed by simulation.  In the classic experiment by
\citeA{Shepard1971}, participants viewed images of three-dimensional
objects and had to determine whether the images depicted the same
object (which differed by a rotation) or two separate objects (which
differed by a reflection and a rotation). They found that people's
response times had a strong linear correlation with the minimum angle
of rotation, a result which led to the conclusion that people solve
this task by ``mentally rotating'' the objects until they are
congruent.  This explanation leaves several questions unanswered,
however. How do people know the axis around which to rotate the
objects? If the axis is known, how do people know which direction to
rotate the objects?  And finally, how do people know how long to
rotate?

@@

Previous models of mental rotation have largely focused on the
representation of mental images, rather than how people decide
\textit{which} mental images to generate. \citeA{Kosslyn:1977tv}
proposed a model of the mental imagery buffer, but not the planning
process of how it is used. Similarly, \citeA{Julstrom:1985va} and
\citeA{Glasgow:1992tj} were mostly concerned with modeling the
representational format underlying imagery. Although
\citeA{Anderson1978} emphasized the importance of considering both
representation and process, he dismissed the problem of determining
the direction of rotation as a ``technical difficulty''.

@@

If people are not computing the rotation beforehand, what might they
be doing? To answer this question, we perform a rational analysis of
the problem of mental rotation
\cite{Marr:1983to,anderson90,Shepard:1987tt}. At the computational
level, we can say that the \textit{problem} is to determine which
spatial transformations an object has undergone based on two images of
that object (which do not include information about point
correspondences). At the algorithmic level, we are constrained by the
notion that mental images must be transformed in an analog manner (or
in a way that is at least approximately analog), and that mental
images are time-consuming and effortful to generate. Thus, the
\textit{goal} is to make this determination while performing a minimum
amount of computation (i.e., as few rotations as possible).

The original ``congruency'' hypothesis \cite{Shepard1971} actually is
a rational solution to this problem, in the sense that the smallest
amount of computation coincides with rotating through the minimum
angle. However, it violates the constraint that we do not know the
points of correspondence between the images, which is exactly what
necessitates the use of imagery. Noting that a rational solution need
not necessarily maintain a single trajectory of rotation, we explore
an alternative rational model, which--rather than computing the angle
of rotation ahead of time--engages in an \textit{active sampling}
strategy.

Active sampling is the idea that when initial information is
incomplete, people will gather new information in a manner that
increases certainty about the problem space. This notion has recently
gained support in several areas of cognitive science
\cite<e.g.,>{Nelson2007,Gureckis:2012gu}, including other spatial
domains \cite{Juni:2011vo}. In the case of mental rotation, if people
do not know the minimum angle beforehand, then actively performing
rotations may be the best way to gather evidence about the similarity
between the observed shapes.

@@

In this section, we formalize our rational analysis and propose four
models of mental rotation: one based on existing models; two which are
extensions of the first but with relaxed assumptions; and one based on
the active sampling approach.

@@

Formally, we denote the shapes as $X_a$ and $X_b$ and assume $X_b$ is
generated by a transformation of $X_a$, i.e. $X_b=f(X_a, \theta, h)$,
where $\theta$ is a rotation, $\hi$ is the hypothesis that the images
depict the same object, and $\hf$ is the hypothesis that the images
depict mirror-image objects. The posterior probability of each
hypothesis given the observed shapes is then:
\begin{equation}
  p(h\ \vert\ X_a, X_b) \propto \int p(X_b\ \vert\ X_a, \theta, h)p(X_a)p(h)p(\theta)\dif\theta,
  \label{eq:posterior}
\end{equation}
where $p(X_b\ \vert\ X_a, \theta, h)$ is the probability that $X_b$
was generated from $X_a$, and $p(X_a)$ is the probability of observing
$X_a$. Because we ultimately want to determine which hypothesis is
more likely, the quantity of interest is a ratio $\ell:=p(\hi\ \vert\
X_a, X_b) / p(\hf\ \vert\ X_a, X_b)$ which (assuming that all
rotations are equally likely) is equivalent to:
\begin{equation}
  \ell = \frac{\left(\int p(X_b\ \vert\ X_a, \theta, \hi)\dif\theta\right)p(\hi)}{\left(\int p(X_b\ \vert\ X_a, \theta, \hf)\dif\theta\right)p(\hf)}.
  \label{eq:lh-ratio}
\end{equation}
If $\ell > 1$, then we accept the hypothesis that the images depict
the same object ($\hi$); if $\ell < 1$, then we accept the hypothesis
that the images depict flipped objects ($\hf$).

@@

\begin{figure*}[t]
  \begin{center}
    \includegraphics[width=\textwidth]{../../figures/D/model-traces.pdf}
    \caption{\textbf{Example stimuli and model behavior.} On the left
      are two stimulus pairs, one ``same'' (top) and one ``flipped''
      (bottom), both with a rotation of $120^\circ$. The grey plots
      show the actions that each model took for these stimuli. The
      color of each point (action) indicates whether the model's
      current mental image was generated under the hypothesis that the
      objects are the same (red) or flipped (blue). The similarity
      functions for each of these hypotheses are shown on the right.}
    \label{fig:model-traces}
  \end{center}
\end{figure*}

@@

One hypothesis is that people compute the direction and extent of
rotation beforehand using \textit{a prior} knowledge of the
correspondence between points in the images
\cite{Funt:1983wn,Just:1985uu}.  To reflect this hypothesis, we
created an ``oracle'' model which is told which points on each shape
correspond. From that correspondence, it computes the correct rotation
and rotates through it.

@@

What if the point correspondences are unknown? An intuitive and simple
model which obeys this constraint first rotate in the direction that
increases similarity, and then stop rotating once a ``match'' is
found, as determined by a threshold on the value of $S$. If all
rotations have been exhausted, then flip, and try rotating again. We
assume that the locations where $S$ is greater than the threshold
correspond to the true $\theta$ (or points near the true
$\theta$). So, as with the \Oc{} model, Eq.~\ref{eq:lh-ratio} becomes
a likelihood ratio.

@@

In the current formulation of the problem, choosing the threshold is
straightforward because we know both the exact geometry of the shapes
and that a linear transformation exists which will perfectly align
them. However, this choice is not always clear \textit{a priori}, as
the global optimum depends on many factors (e.g., shape complexity,
dimensionality, perceptual uncertainty, and whether the shapes are
identical).  One way to deal with the problem of choosing a threshold
would simply be to perform an exhaustive search or use a global
optimization strategy. However, these strategies would not result in
the linear response time found by \citeA{Shepard1971}. A second
alternative, which could potentially demonstrate the linear response
time result, might instead just perform a Hill Climbing (\Hc{}) search
for a \textit{local} maximum. As with the \Oc{} and \Th{} models,
Eq.~\ref{eq:lh-ratio} becomes a generalized likelihood ratio test.

@@

While the previous few models have all focused on \textit{searching}
for the global maximum, the goal is really to just approximate
Eq.~\ref{eq:lh-ratio} while still obeying the constraint of sequential
rotations. Based on evidence that people use ``active sampling''
\cite<e.g.,>{Nelson2007,Gureckis:2012gu}, we hypothesize a model that
actively monitors the usefulness of the rotations it is
performing. Instead of searching for a maximum, we maintain a
probability distribution over our \textit{estimate} of
Eq.~\ref{eq:lh-ratio}, and then sample actions which are expected to
improve that estimate.  This strategy has several benefits. First, it
does not make assumptions about the shape or scale of the similarity
function. Second, by choosing to sample places which are informative,
this method implicitly minimizes the amount of rotation.

Formally, we denote $Z_h$ as our estimate of the likelihood for
hypothesis $h$, and write its distribution as:
\begin{equation}
  p(Z_h) = \int \left[\int S(\Xb, f(\Xa, \theta, h))p(\theta)\dif\theta\right] p(S)\dif S,
\end{equation}
where $S$ is the similarity function, and $p(S)$ is a prior over
similarity functions.  This method of estimating an integral is known
in the machine-learning literature as \textit{Bayesian Quadrature}
\cite{Diaconis:1988uo,OHagan:1991tx,Osborne:2012tm}, or BQ.

Denoting $S_h=S(\Xb, f(\Xa, \theta, h))$, we first place a
\textit{Gaussian Process} \cite{Rasmussen:2006vz}, or GP, prior on the
log of $S_h$ in order to enforce positivity after it is exponentiated,
i.e. $\mathbb{E}[Z_h] \approx \int
\exp(\mu_h(\theta))p(\theta)\dif\theta$, where $\mu_h:=\mu(\log S_h)$
is the mean of the log-GP \cite{Osborne:2012tm}.  However, this
integral is still intractable. To approximate it, we fit a second GP
over points sampled from the log-GP, which we denote as
$\bar{S}_h:=\exp(\mu_h)$.\footnote{Note that the approach used here
  differs slightly from the one presented in \citeA{Osborne:2012tm}
  and is based on personal communication with one of the authors
  \cite{Duvenaud:2013td}.} Then, from \citeA{Duvenaud:2013td}:
\begin{align}
  \mathbb{E}[Z_h] &\approx \int \bar{\mu}_h(\theta)p(\theta)\dif\theta,\\
  \mathbb{V}(Z_h) &\approx \iint \mathrm{Cov}_h(\theta,
  \theta^\circ)\bar{\mu}_h(\theta)\bar{\mu}_h(\theta^\circ)p(\theta)p(\theta^\circ)\dif\theta\dif\theta^\circ,
\end{align}
where $\bar{\mu}_h:=\mu(\bar{S}_h)$ is the mean of the second GP, and
$\mathrm{Cov}_h:=\mathrm{Cov}(\log S_h)$ is the covariance of the
log-GP.

@@

We randomly generated 20 shapes of five or six vertices (e.g.,
Fig.~\ref{fig:model-traces}). For each shape, we computed 20 ``same''
and 20 ``flipped'' stimuli pairs, with 18 rotations ($\theta$) spaced
at $20^\circ$ increments between $0^\circ$ and $360^\circ$ (with
$0^\circ$ and $180^\circ$ repeated twice, in order to gather an equal
number of responses for each angle between $0^\circ$ and
$180^\circ$). ``Same'' pairs were created by rotating $\Xa$ by
$\theta$; ``flipped'' pairs were first reflected $\Xa$ across the
$y$-axis, then rotated by $\theta$.

We generated five additional shapes to be used in a practice block of
10 trials. Across these trials, there was one ``flipped'' and one
``same'' repetition of each shape and each angle ($60^\circ$,
$120^\circ$, $180^\circ$, $240^\circ$, or $300^\circ$) such that no
shape was presented at the same angle twice. We also generated a sixth
shape to include with the instructions.  This shape had both a
``flipped'' and ``same'' version, each rotated to $320^\circ$.

@@

All participants saw the same 10 practice trials as described
above. There were 720 unique experimental stimuli (20 shapes $\times$
18 angles $\times$ 2 reflections), though because stimuli with
rotations of $0^\circ$ or $180^\circ$ were repeated twice, there were
800 total experimental stimuli. These stimuli were split across eight
conditions in the following manner: first, stimuli were split into
four blocks of 200 trials. Within each block, each shape was repeated
ten times and each rotation was repeated ten times (five ``same'',
five ``flipped''), such that across all blocks, each stimulus appeared
exactly once. Each block was then split in half. Participants
completed two half-blocks (not necessarily both from the original full
block).

@@

Out of the 247 participants, 200 (81\%) were included in our
analyses. Of the rest, we excluded 10 (4\%) because of an experimental
error, 6 (2.4\%) because they had already completed a related
experiment, and 31 (12.6\%) because they failed a comprehension
check. Comprehension was defined as correctly answering at least 85\%
of stimuli with a minimum rotation of $0^\circ$ or $20^\circ$.
Additionally, we excluded 82 trials for which the response time was
either less than 100 milliseconds or greater than 20 seconds.

For each model, we ran 50 samples for each of the 800 experimental
stimuli. The step size parameter ($\sigma_\epsilon$) was fit to human
response times for each of the models, resulting in
$\sigma_\epsilon=0.6$ for the \Th{} and \Bq{} models and
$\sigma_\epsilon=0.1$ for the \Oc{} and \Hc{} models. We also ran the
models under two different priors, $p(h=0)=0.5$ (the ``equal'' prior)
and $p(h=0)=0.55$ (the ``unequal'' prior). This only had an effect on
the stopping criteria for the \Bq{} model.\footnote{The prior also had
  a small effect on the accuracy of the \Hc{} model, because sometimes
  the local maxima were small enough that Eq.~\ref{eq:mle-lh-ratio}
  became dominated by the prior.}

@@

For analyses of response time, confidence intervals around harmonic
means were computed using a bootstrap analysis using 10000 bootstrap
samples (sampled with replacement). Response times were only
considered for correct responses.  We also used a bootstrap analysis
of 10000 bootstrap samples to compute the confidence intervals around
both Spearman ($r_s$) and Pearson ($r$) correlations.  Unless
otherwise specified, all correlations were computed over 720
datapoints (one for each stimulus). For analyses of accuracy,
confidence intervals were computed from a binomial proportion with a
Jeffrey's beta prior.  To test if participants or a model was above
chance on a particular stimulus, we used the same binomial proportion,
and simply tested whether $p>0.05$ for the 50\% mark.

\TODO{need to recompute correlations}

@@

The average accuracy across all stimuli was \ExpAccuracy{}, though
there were \ExpNumChance{} stimuli (out of 720) for which people were
not above chance.  The minimum angle was also correlated with
participants' average per-stimulus accuracy, as shown in
Fig.~\ref{fig:theta-accuracy-corrs}, though significantly more so for
``same'' pairs. This is the same result found both by
\citeA{Cooper:1975wp} and \citeA{Gardony:2013gn}.

@@

The number of actions taken by the \Th{} model was significantly
correlated with the minimum angle (Fig.~\ref{fig:theta-time-corrs}),
though this effect was less strong for individual shapes (e.g.,
Fig.~\ref{fig:response-time-stimulus}).  The \Th{} model was able to
explain a moderate amount of the variance in human response times,
with a correlation of \ThTimeCorr{}
(Fig.~\ref{fig:human-model-corrs}). Like the \Oc{} model, the \Th{}
model had 100\% accuracy, and thus there was no correlation between
minimum angle and accuracy.

As noted above, we fit $\sigma_\epsilon=0.6$ for the \Th{} model. This
has the interesting effect of actually causing the \Th{} to
\textit{over}rotate (see, for example, the ``flipped'' pair trace in
Fig.~\ref{fig:model-traces}). This is because the step size is large
enough that it sometimes misses the global maximum, and must do
another full rotation to find it. However, when the step size is small
enough to always find the maximum, the fit to human data is slightly
worse.

@@

The \Bq{} model was quite accurate overall (equal prior:
(\BqAccuracy{}; unequal prior: \BqpAccuracy{}). With the equal prior,
there was \BqNumChance{} stimulus for which it was not above chance;
with the unequal prior, there were \BqpNumChance{}. There was a
significant effect of rotation on the \Bq{} model's accuracy
(Fig.~\ref{fig:theta-accuracy-corrs}), and it explained about the same
amount of variance in people's accuracy as the \Hc{} model (equal
prior: \BqAccuracyCorr{}; unequal prior: \BqpAccuracyCorr{}).

Because the \Bq{} model relies on Eq.~\ref{eq:lh-ratio} for its
stopping criteria (as opposed to just finding a maximum), the prior
$p(h)$ has a fairly significant effect, as shown in
Fig.~\ref{fig:response-time-accuracy}. With just a small bias of
$p(h=0)=0.55$, there is a clear separation in response times for
``same'' versus ``flipped'' stimuli; this separation is similar to the
trend also observed in human response times. While the prior also had
an effect on accuracy, this effect did not reflect the trends in human
behavior.
